---
title: "Beta Regression in Stan"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Beta Regression in Stan

This is code to implement beta regression similar to the [`betareg` package](https://cran.r-project.org/web/packages/betareg/index.html) in [Stan](http://mc-stan.org/).

## Simulated data

A minimal simulation model can be used to relate the value of a beta distributed random variable, `y`, to two covariates, `x1` and `x2`.

```{r, ggplot2, message = F, warning = F}
library(betareg)
library(rstan)
library(dplyr)

N = 500
x1 = rnorm(N)
x2 = rnorm(N)
X = cbind(1, x1, x2)
beta = c(-1,.2,-.3)
mu = plogis(X%*%beta)  # add noise if desired + rnorm(N, sd=.01)
phi = 10
A = mu*phi
B = (1-mu)*phi
y = rbeta(N, A, B)
qplot(y, geom = "histogram", binwidth = .025)

```

## betareg

The simulated data can be fit using the `betareg` package as follow:

```{r}
brmod <- betareg(y ~ ., data = data.frame(y, X[,-1]))
summary(brmod)
```

## Stan

We can fit the same model Stan with the following code:

```{r, message = F, warning = F, results = "hide"}

stan_beta <- "
data {
  int<lower=1> N;
  int<lower=1> K;
  vector<lower=0,upper=1>[N] y;
  matrix[N,K] X;

}

parameters {
  vector[K] gamma;
  real<lower=0> phi;
}
transformed parameters{
  vector<lower=0,upper=1>[N] mu;    // transformed linear predictor
  vector<lower=0>[N] A;             // parameter for beta distn
  vector<lower=0>[N] B;             // parameter for beta distn

  for (i in 1:N)
    mu[i] = inv_logit(X[i,] * gamma);   
  A = mu * phi;
  B = (1.0 - mu) * phi;
}
model {

  // priors
  // gamma ~ normal(0, 1);   
  // phi ~ cauchy(0, 5);               // different options for phi  
  // phi ~ inv_gamma(.001, .001);
  // phi ~ uniform(0, 500);          // put upper on phi if using this
  // likelihood
  y ~ beta(A, B);

}
"
# Stan data list
dat = list(N = length(y), K=ncol(X), y=y, X=X)

beta_stan_test <- stan(model_code = stan_beta,
     data       = dat,
     pars       = c("gamma", "phi"))
```


Which yields the following output:
```{r}
summary(beta_stan_test)$summary
```
