---
title: "Beta Regression in Stan"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Beta Regression in Stan

This is code to implement beta regression similar to the [`betareg` package](https://cran.r-project.org/web/packages/betareg/index.html) in [Stan](http://mc-stan.org/).

## Simulated data

A minimal simulation model can be used to relate the value of a beta distributed random variable, `y`, to two covariates, `x1` and `x2`.

```{r, ggplot2, message = F, warning = F}
library(betareg)
library(rstan)
library(dplyr)


N = 500
x1 = rnorm(N)
x2 = rnorm(N)
X = cbind(1, x1, x2)
beta = c(-1,.2,-.3)
gamma = c(2, .25, -.5)
mu = plogis(X %*% beta)  # add noise if desired + rnorm(N, sd=.01)
phi = exp(X %*% gamma)
A = mu*phi
B = (1-mu)*phi
y = rbeta(N, A, B)

qplot(y, geom = "histogram", binwidth = .025)

```

## betareg

The simulated data can be fit using the `betareg` package as follow:

```{r}
brmod <- betareg(y ~ x1 + x2 | x1 + x2, data = data.frame(y, X[,-1]))
summary(brmod)
```

## Stan

We can fit the same model Stan with the following code:

```{r, message = F, warning = F, results = "hide"}

stan_beta <- "
data {
  int<lower=1> N;
  int<lower=1> K;
  int<lower=1> J;
  vector<lower=0,upper=1>[N] y;
  matrix[N,K] X;
  matrix[N,J] Z;
}

parameters {
  vector[K] beta;
  vector[J] gamma;
}

transformed parameters{
  vector<lower=0,upper=1>[N] mu;    // transformed linear predictor for mean of beta distribution
  vector<lower=0>[N] phi;           // transformed linear predictor for precision of beta distribution
  vector<lower=0>[N] A;             // parameter for beta distn
  vector<lower=0>[N] B;             // parameter for beta distn

  for (i in 1:N) {
    mu[i]  = inv_logit(X[i,] * beta);   
    phi[i] = exp(Z[i,] * gamma);
  }

  A = mu .* phi;
  B = (1.0 - mu) .* phi;
}

model {
  // priors

  // likelihood
  y ~ beta(A, B);
}

generated quantities{
  vector[N] log_lik;
  vector[N] log_lik_rep;
  vector<lower=0,upper=1>[N] y_rep;
  real total_log_lik;
  real total_log_lik_rep;
  
  int<lower=0, upper=1> p_omni;

  for (n in 1:N) {
    log_lik[n] = beta_lpdf(y[n] | A[n], B[n]);
    y_rep[n] = beta_rng(A[n], B[n]);
    log_lik_rep[n] = beta_lpdf(y_rep[n] | A[n], B[n]);
  }

  total_log_lik = sum(log_lik);
  total_log_lik_rep = sum(log_lik_rep);

  p_omni = (total_log_lik_rep > total_log_lik);
}
"

# Stan data list
dat = list(N = length(y), 
            K = dim(X)[2], 
            J = dim(X)[2], 
            y = y, 
            X = X, 
            Z = X)

beta_stan_test <- stan(model_code = stan_beta,
     data       = dat,
     pars       = c("beta", "gamma"))
```


Which yields the following output:
```{r}
summary(beta_stan_test)$summary
```
